---
title: "LITAP"
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/FRDC-SHL/LITAP/workflows/R-CMD-check/badge.svg)](https://github.com/FRDC-SHL/LITAP/actions)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 15, fig.height = 10, eval = TRUE)
```

__Landscape Integrated Terrain Analysis Package (LITAP)__

LITAP is a software package aimed at providing open access code for terrain
analysis and landscape and hydrology models built on terrain attributes. A major
component of LITAP is founded on R. A. (Bob) MacMillan's LandMapR suite of
programs for flow topology and landform segmentation analyses with extended new
parameters and methodologies developed by Drs. Sheng Li and Steffi LaZerte.
Another component of LITAP focuses on the calculations and uses of directional
terrain attributes developed by Dr. Sheng Li. This project is actively under
development and we plan to add more features to support modelling of surficial
hydrology, soil erosion, landscape zoning, hydrologic response units and buffer
zone delineation.

LITAP is a project funded by Agriculture and Agri-Food Canada and
directed by [Dr. Sheng Li](https://profils-profiles.science.gc.ca/en/profile/sheng-li-phd).

See articles for new features specific to LITAP (coming soon)!

Currently LITAP includes major functions

- `flow_mapper()` - based on FlowMapR
- `form_mapper()` - based on FormMapR
- `facet_mapper()` - based on FacetMapR
- `wepp_mapper()` - based on WeppMapR

See the companion website for more details: <https://FRDC-SHL.github.io/LITAP/>

## Installation

Use the `remotes` package to directly install R packages from github:

```{r, eval = FALSE}
install.packages("remotes") # If not already installed
remotes::install_github("FRDC-SHL/LITAP") 
```

## Example files

LITAP comes with several example files (`testELEV.dbf`, `testELEV_mini.dbf`,
`arule.dbf`, and `crule.dbf`). These are found in the "extdata" folder inside
the LITAP package folder. If you are unsure where this folder is, use the
function `system.file()`:

```{r, eval = FALSE}
system.file("extdata", package = "LITAP")
```

Now you can copy and paste these files to your working folder to try out the following examples.

```{r, include = FALSE}
file.copy(system.file("extdata", "testELEV.dbf", package = "LITAP"), ".")
file.copy(system.file("extdata", "arule.dbf", package = "LITAP"), ".")
file.copy(system.file("extdata", "crule.dbf", package = "LITAP"), ".")
```

## Basic Usage: `flow_mapper()`

Load the package:
```{r}
library(LITAP)
```

First, specify the dem file and the number of rows and columns:

```{r, eval = FALSE}
flow_mapper(file = "testELEV.dbf", nrow = 90, ncol = 90, grid = 5)
```

Can also specify pit removal parameters:

```{r, eval = FALSE}
flow_mapper(file = "testELEV.dbf", nrow = 90, ncol = 90, grid = 5, 
             max_area = 5, max_depth = 0.2)
```

As well as the location of output files:
```{r}
flow_mapper(file = "testELEV.dbf", nrow = 90, ncol = 90, grid = 5, out_folder = "./Output/")
```

## Basic Usage: `form_mapper()`

`form_mapper()` uses output from `flow_mapper()` and requires a grid size.

```{r}
form_mapper(folder = "./Output/")
```

Optionally, users can also define channels and ridges according to the number of up-/down-slope cells that flow through the cell in question.

```{r, eval = FALSE}
form_mapper(folder = "./Output/", str_val = 10000, ridge_val = 10000)
```

## Basic Usage: `facet_mapper()`

`facet_mapper()` uses output from `flow_mapper()` and `form_mapper()` and requires a CRULE file
and, optionally, an ARULE file.

```{r}
facet_mapper(folder = "./Output/", crule = "crule.dbf", arule = "arule.dbf")
```

If an ARULE file is not provided, LITAP will derive the rules based on the input 
files (as in Li et al. 2011, Canadian Journal of Soil Science 91(2), 251-266). 
The derived A rules will be output to "afile_derived.csv".

```{r, eval = FALSE}
facet_mapper(folder = "./Output/", crule = "crule.dbf")
```

## Basic Usage: `wepp_mapper()`

`wepp_mapper()` uses output from `flow_mapper()`

```{r, eval = FALSE}
wepp_mapper(folder = "./Output/")
```

Optionally, users can also define the maximum length of channel cells 
(after which they will be split into separate segments), as well as the 
upslope threshold to define channel cells 

```{r, eval = FALSE}
wepp_mapper(folder = "./Output/", chan_length = 500, upslope_threshold = 500)
```

## Multiple file types

LITAP accepts multiple file types and can automatically assess the number of rows and columns, depending on the type (see `?load_file` for details and requirements):

```{r, eval = FALSE}
flow_mapper(file = "testELEV.csv")
flow_mapper(file = "testELEV.grd")
flow_mapper(file = "testELEV.flt")
```

## Output

Output files can be .csv or .rds (R data files) in the output folders "flow", "form", "facet" and "wepp". Additionally, an html report summarizing the `flow_mapper()` run is included in the output folder (`testELEV_final_report.html`).

See the [LITAP website](http://FRDC-SHL.github.io/LITAP/) for more details and examples

```{r, include = FALSE}
unlink("./testELEV", recursive = TRUE)
unlink("./Output", recursive = TRUE)
unlink(c("./testELEV.dbf", "arule.dbf", "crule.dbf"))
```
