---
title: "README"
author: Steffi LaZerte
---

```{r, include = FALSE}
knitr::opts_chunk$set(fig.width = 15, fig.height = 10, tidy = TRUE, eval = FALSE)
```

# LITAP

__Landscape Integrated Terrain Analysis Package (LITAP)__ for the analysis of waterflow based on elevation data and pit removals. Based on R. A. (Bob) MacMillan's FlowMapR program.

See the companion website for more details: <http://steffilazerte.github.io/LITAP>

## Installation

Use the `devtools` package to directly install R packages from github:

```{r}
install.packages("devtools") # If not already installed
devtools::install_github("steffilazerte/LITAP") 
```

## Basic Usage

Must specify the dem file and the number of rows and columns:

```{r}
complete_run(file = "testElev.dem", nrow = 100, ncol = 100)
```

Can also specify pit removal parameters:

```{r}
complete_run(file = "testElev.dem", nrow = 100, ncol = 100, 
             max_area = 5, max_depth = 0.2)
```

As well as the location of output files:
```{r}
complete_run(file = "testElev.dem", nrow = 100, ncol = 100, folder_out = "./Output/")
```

## Output

Output files include .rds (R data files) and .csv files in the "Final" folder, .dbf files in the "dbf" folder and backup files (for resuming runs) in teh "Backup" folder. Additionally, an html report summaring the run is included in the output folder

The .dbf files are designed to imitate the output of the original FlowMapR program (note that the .rds and the .csv files do not).

See the [LITAP website](http://steffilazerte.github.io/LITAP/) hosted on github for more details and examples
