---
title: "LITAP Final Report"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float:
      collapsed: false
params: 
   file: ""
   out_files: ""
   run: ""
   max_area: ""
   max_depth: ""
   nrow: ""
   ncol: ""
   rlim: ""
   clim: ""
---
<style>
.figure {
  max-width: 100%;
  margin: auto;
}
</style>


```{r load_files, include = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 10, tidy = TRUE, echo = FALSE)

if(is.null(params$rlim)) rlim <- "none" else rlim <- params$rlim
if(is.null(params$clim)) clim <- "none" else clim <- params$clim

files <- params$out_files

library(LITAP)
library(gridExtra)

cat(files)
if(length(stringr::str_subset(files, paste0(run, "_dem_fill.rds"))) > 0) dem <- readRDS(stringr::str_subset(files, paste0(run, "_dem_fill.rds")))
if(length(stringr::str_subset(files, paste0(run, "_pit.rds"))) > 0) pit <- readRDS(stringr::str_subset(files, paste0(run, "_pit.rds"))) #cat(stringr::str_subset(files, paste0(run, "_pit.rds"))) 
if(length(stringr::str_subset(files, paste0(run, "_pond.rds"))) > 0) pond <- readRDS(stringr::str_subset(files, paste0(run, "_pond.rds")))
if(length(stringr::str_subset(files, paste0(run, "_fill.rds"))) > 0) fill <- readRDS(stringr::str_subset(files, paste0(run, "_fill.rds")))
```

## Details

**Input File**:   

`r params$file`  

__Parameters:__
 
 - max_area: `r params$max_area`
 - max_depth: `r params$max_depth`
 - nrow: `r params$nrow`
 - ncol: `r params$ncol`
 - rlim: `r rlim`
 - clim: `r clim`
 
__LITAP version:__ 

`r packageVersion("LITAP")`



## Local files output (.csv)

```{r list_files, results = "asis"}
lf <- stringr::str_subset(files, "csv$")
cat(paste0("- [", basename(lf), "](", lf, ")\n"))
```


## Maps
```{r maps, fig.height = 7}
grid.arrange(flow_plot(db = dem, type = "relief") + labs(title = "relief"),
             flow_plot(db = dem, type = "elevation") + labs(title = "elevation"),
             nrow = 1)
```

- Initial watersheds represent watersheds identified in the very first pass
- Local watersheds represent watersheds after the initial pit removal has been performed
- Pond watersheds represent how the local watersheds overflow into various pond sheds

```{r flow_plots}
grid.arrange(flow_plot(db = dem, type = "elevation", pits = TRUE, shed = TRUE, shed_type = "initial") + 
               labs(title = "Initial Watersheds") +
               theme(plot.margin = unit(c(0, 0, 5 ,0), "mm")),
             flow_plot(db = dem, type = "elevation", pits = TRUE, shed = TRUE, shed_type = "local") + 
               labs(title = "Local Watersheds")+
               theme(plot.margin = unit(c(0, 0, 5 ,0), "mm")),
             flow_plot(db = dem, type = "elevation", pits = TRUE, shed = TRUE, shed_type = "pond") + 
               labs(title = "Pond Watersheds")+
               theme(plot.margin = unit(c(5, 0, 0 ,0), "mm")),
             flow_plot(db = dem, type = "elevation", pits = TRUE, shed = TRUE, shed_type = "fill") + 
               labs(title = "Fill Watersheds")+
               theme(plot.margin = unit(c(5, 0, 0 ,0), "mm")),
             ncol = 2)
```

## Results

__NOTE:__  

- NextPit values do not match the original values from LandMapR
- Varatio/Pit Vol/Pit Area may be slightly different as well

### Watershed stats - Pit Table
```{r pit_table}
if(exists("pit")) DT::datatable(pit, rownames = FALSE) else message("No pit table")
```

### Watershed stats - Pond Table
```{r pond_table}
if(exists("pond")) DT::datatable(pond, rownames = FALSE) else message("No pit table")
```

### Watershed stats - Fill Table
```{r fill_table}
if(exists("fill")) DT::datatable(fill, rownames = FALSE, filter = "none", options = list(ordering=F)) else message("No fill table")
```
