---
title: "Test - Hypothetical Hummocky Low Relief"
author: "Steffi LaZerte"
date: "`r format(Sys.Date())`"
output: html_document
---

<style>
.figure {
  max-width: 50%;
  margin: auto;
}
.table {
  width: auto;
  margin: auto;
}
</style>

```{r, include = FALSE}

knitr::opts_chunk$set(fig.width = 12, tidy = TRUE)

library(tidyverse)
library(gridExtra)
library(LITAP)

o <- "~/Projects/Business/LandmapR/TestFiles/01_H1/FlowMapR/"
n <- "~/Projects/Business/LandmapR/Runs/H10/"
run <- "H10"


# Original Run DEM
orig_dem_plot <- load_orig(file = paste0(o, run, "dem.dbf"))
orig_dem <- load_odem(file = paste0(o, run, "dem.dbf"))

# LITAP Runs - DEM FILES
pond_dem <- load_new(file = paste0(n, "/final/", run, "_dem_pond.rds"))
fill_dem <- load_new(file = paste0(n, "/final/",run, "_dem_fill.rds"))
dem <- load_ndem2(file = paste0(n, "/dbf/", run, "_dem.dbf"))

# Original Run - SHED STATS
orig_pond <- load_ostat(file = paste0(o, run, "pond.dbf"))
orig_fill <- load_ostat(file = paste0(o, run, "fill.dbf"))
orig_pit <- load_ostat(file = paste0(o, run, "pit.dbf"))

# LITAP RUNS - SHED STATS
initial <- load_new(file = paste0(n, "/final/", run, "_initial.rds"))
local <- load_ndem1(file = paste0(n, "/dbf/", run, "_local.dbf"))
pond <- load_ndem1(file = paste0(n, "/dbf/", run, "_pond.dbf"))
fill <- load_ndem1(file = paste0(n, "/dbf/", run, "_fill.dbf"))
pit <- load_ndem1(file = paste0(n, "/dbf/", run, "_pit.dbf"))

# Original run - Inverted pits
orig_ipit <- load_ostat(file = paste0(o, run, "ipit.dbf"))

# LITAP RUNS - Inverted pits
ipit <- load_ndem1(file = paste0(n, "/dbf/", run, "_ipit.dbf"))

idem <- load_new(file = paste0(n, "/final/", run, "_dem_ilocal.rds"))
orig_idem <- load_orig(file = paste0(o, run, "idem.dbf"))

```

# Details

2.22 min to run


#### Original Dem (Relief)
```{r, echo = FALSE, fig.width = 5}
flow_plot(orig_dem_plot)
```

#### Original Dem (Elevation)
```{r, echo = FALSE, fig.width = 5}
flow_plot(orig_dem_plot, type = "elevation")
```


# Visual Comparison


## Original vs. New Local Shed assignment

- The first shed shows initial pits which are removed in the first step of pit removal. This dem doesn't exist in the original program.
- The next two are the local sheds (shedno) after amalgamating the small ones
- Sheds are identical

```{r, echo = FALSE}
grid.arrange(
  flow_plot(fill_dem, shed = TRUE, shed_type = "initial") +
    labs(title = "Initial (New)"),
  flow_plot(orig_dem_plot, shed = TRUE, shed_type = "local") +
    labs(title = "Original"),
  flow_plot(fill_dem, shed = TRUE, shed_type = "local") +
    labs(title = "New"), nrow = 1)
```

## Original vs. New Fill Shed assignment

- These are the fill sheds, which reflect how sheds combined during the third pit removal process
- Shapes should be identical to pond filling, just the numbers may be different
- Sheds are identical

```{r, echo = FALSE}
grid.arrange(
  flow_plot(orig_dem_plot, shed = TRUE, shed_type = "fill") +
    labs(title = "Original"),
  flow_plot(fill_dem, shed = TRUE, shed_type = "fill") +
    labs(title = "New"), nrow = 1)
```

# Numeric Comparison

## Compare pit stats

__Original Pit data__

```{r, echo = FALSE}
DT::datatable(orig_pit)
```

__New Pit data__

```{r, echo = FALSE}
DT::datatable(pit)
```

__Differences__

```{r}
c <- compare(pit, orig_pit)
```

- For some reason original End pit, Pit Vol and Varatio, are different for last two pits:

```{r}
pit[c$EndPit,] %>% select(ShedNo, EndPit, ShedArea, PitArea, PitVol, Varatio, Becomes)
orig_pit[c$EndPit,] %>% select(ShedNo, EndPit, ShedArea, PitArea, PitVol, Varatio, Becomes)
```

- Pit area off, probably likely due to tie breaking which is more of a problem when the watershed is uniform
```{r}
pit[c$PitArea,] %>% select(ShedNo, EndPit, ShedArea, PitArea, PitVol, Varatio, Becomes)
orig_pit[c$PitArea,] %>% select(ShedNo, EndPit, ShedArea, PitArea, PitVol, Varatio, Becomes)
```

- Drains to off?
```{r}
pit[c$DrainsTo,] %>% select(ShedNo, OutRec, DrainsTo, Becomes)
orig_pit[c$DrainsTo,] %>% select(ShedNo, OutRec, DrainsTo, Becomes)
```


## Compare pond stats

__Original Pond data__

```{r, echo = FALSE}
DT::datatable(orig_pond)
```
 __New Pond data__
```{r, echo = FALSE}
DT::datatable(pond)
```

Cells are so regular that tie breaking caused different flows paths between the two programs

For example, in the original file, watershed 16 drains to 13 (south), in LITAP watershed 16 drains to 12 (west)

There is no difference between the pour points, so it would just be a matter of which one was visited first.

```{r}
pond %>% filter(ShedNo == 16) %>% select(ShedNo, PitRec, PourElev, DrainsTo, OutElev)
orig_pond %>% filter(ShedNo == 16) %>% select(ShedNo, PitRec, PourElev, DrainsTo, OutElev)
```

However the end product is the same (see pit files above)


```{r}
pond <- pond %>% arrange(ShedNo)
orig_pond <- orig_pond %>% arrange(ShedNo)
```


## Compare fill stats

__Original Fill data__

```{r, echo = FALSE}
DT::datatable(orig_fill)
```

__New Fill data__

```{r, echo = FALSE}
DT::datatable(fill)
```

Order of pit filling is the same 

__Differences__

```{r}
fill <- arrange(fill, ShedNo)
orig_fill <- arrange(orig_fill, ShedNo)
c <- compare(fill, orig_fill)
```

- All important values (pit area, pit vol, varatio, becomes) are identical, or only different by a small degree
- Out sheds (DrainsTo) not all the same... does this matter?
- EndPit is different but not clear why original program has different end pits... overlooked?


## Inverted DEM
Black dots indicate pits, or, as this is inverted, peaks

```{r, echo = FALSE}
grid.arrange(
  flow_plot(orig_idem, shed = TRUE, shed_type = "fill", pits = TRUE) +
    labs(title = "Original"),
  flow_plot(idem, shed = TRUE, shed_type = "local", pits = TRUE) +
    labs(title = "New"), nrow = 1)
```

```{r}
ipit <- arrange(ipit, ShedNo)
orig_ipit <- arrange(orig_ipit, ShedNo)
c <- compare(ipit, orig_ipit)
```

- Pit areas off by a small amount
- Small differences in exact pit location (off by one cell at most)
- Drains To same problem as before with 0
